Kindly Bent to Free Us

### 2019-07-10 kindly/lang.tex
regions

    %   (\kaff_n\lk k\lk\kaff_\infty)
    %   &\svar[\MBORROW]{x}{\tau}^n
    %   &\bvar{\borrow[\MBORROW]{x}}{\borrowty[\MBORROW] k{\tau}}
    %   &Mut
    % \end{tabular}
    \begin{mathpar}
      \inferrule{
        \entail{C}{(\BORROW_n\lk k\lk \BORROW_\infty)}\\
        \BORROW\in\left\{\kun;\kaff\right\}
      }{
        \lregion{C}{}
        {\svar[\BORROW]{x}{\tau}^n}
        {\bvar{\borrow[\BORROW]{x}}{\borrowty[\BORROW] k{\tau}}}
      }
    \end{mathpar}
    \caption{Borrowing rules for bindings -- $\lregion{C}{x}{\E}{\E'}$}
    \label{env:rule:borrow}
  \end{minipage}
\end{figure}

@@

The typing for a region (see the {\sc Region} rule in \ref{selectrules:region})
replaces suspended bindings by borrow bindings, typechecks the body
of the region, and finally ensures that the borrow does not leak outside.
This last check is done with integer indices that correspond to the nesting
level of the region. The kind of a borrow is indexed with the level
corresponding to its region and the constraint $\Cleq{\tau}{\klin_{n-1}}$ ensures that
the return type of the region must live at some enclosing, lower level.

As an example, consider the expression $\region{\Sone x\IBORROW}{f(\borrow[\IBORROW]{c})}$
where $c$ is a linear channel.
The first step is to check that
$\svar[\IBORROW]{c}{\text{channel}}$ is in $\E$. We then define $\E'$ which
corresponds to $\E$ where the suspended binding is replaced by a borrow binding.
%
This is done by the borrow relation, noted $\lregion{C}{x}{\E}{\E'}$ and
whose main rules are shown in \cref{selectrules:region}.
In our case, we consider an immutable binding, so we replace the suspended
binding by $\bvar{\borrow[\IBORROW]{c}}{\borrowty[\IBORROW]{k}{\text{channel}}}$.
We must also check that $\entail{C}{(\kun_n\lk k\lk\kun_\infty)}$.
We can then typecheck the body of the region, $f(\borrow[\IBORROW]{c})$ in our example, and check that the borrow made available in the region does not escape.
This is done by the constraint $\Cleq{\tau}{\klin_{n-1}}$, which ensures
that the kind of $\tau$ does not contain anything whose kind is ``above $n$''
(for instance, $\borrowty[\IBORROW]{\kun_n}{\text{channel}}$ would have kind
$\kun_n$).

These remarks lead to an
algorithm for automatic region annotation: start from each borrow, create
a region, and extend it as much as possible while respecting other borrows
and binders.
A complete specification of the algorithm is provided in \cref{regionannot}.
In the rest of this section, we will only consider terms
with fully annotated regions.

>>>>

    %   (\kaff_n\lk k\lk\kaff_\infty)
    %   &\svar[\MBORROW]{x}{\tau}^n
    %   &\bvar{\borrow[\MBORROW]{x}}{\borrowty[\MBORROW] k{\tau}}
    %   &Mut
    % \end{tabular}
    \begin{mathpar}
      \inferrule{
        \entail{C}{(\BORROW_n\lk k) \wedge (k \lk \BORROW_\infty)}\\
        \BORROW\in\left\{\kun,\kaff\right\}
      }{
        \lregion{C}{}
        {\svar[\BORROW]{x}{\tau}^n}
        {\bvar{\borrow[\BORROW]{x}}{\borrowty[\BORROW] k{\tau}}}
      }
    \end{mathpar}
    \caption{Borrowing rules for bindings -- $\lregion{C}{x}{\E}{\E'}$}
    \label{env:rule:borrow}
  \end{minipage}
\end{figure}

@@

The typing for a region (see the {\sc Region} rule in \cref{selectrules:region})
replaces suspended bindings by borrow bindings
(\cref{env:rule:borrow}), typechecks the body 
of the region, and finally ensures that the borrow does not leak outside.
This last check is done with indices that correspond to the nesting
level of the region. The kind $k$ of the borrow is indexed with the level $n$
corresponding to its region, thanks to the constraint $(\BORROW_n\lk
k)$, and the constraint $\Cleq{\tau}{\klin_{n-1}}$ ensures that 
the return type of the region must live at some enclosing, lower level.

As an example, consider the expression $\region{\Sone x\IBORROW}{f(\borrow[\IBORROW]{c})}$
where $c$ is a linear channel in an environment $\E$.
The first step is to check that $\svar[\IBORROW]{c}{\text{channel}}$
is in $\E$.
When entering the region, rule \TirName{Region} imposes
$\lregion{C}{x}{\E}{\E'}$, which defines $\E'$
corresponding to $\E$ where the suspended binding is replaced by the
borrow binding  $\bvar{\borrow[\IBORROW]{c}}{\borrowty[\IBORROW]{k}{\text{channel}}}$.
To constrain the borrow to this region we impose the constraint
$\entail{C}{(\kun_n\lk k) \wedge (\lk\kun_\infty)}$, which affirms
that the borrow is unrestricted, but can only be used in nesting
levels $n$ and higher.  
WIth that setup we typecheck the body of the region,
$f(\borrow[\IBORROW]{c})$.
The constraint $\Cleq{\tau}{\klin_{n-1}}$ ensures
that the kind of $\tau$ does not contain anything whose kind is
``above level $n$'' and thus confines the borrow to the region
(for instance, $\borrowty[\IBORROW]{\kun_n}{\text{c}}$ has kind
$\kun_n$ and thus cannot be returned).

As it would be cumbersome to markup regions manually, we developed an
algorithm that starts from each borrow, creates a region around it,
and extends it as much as possible while respecting other borrows
and binders. The algorithm (see \cref{regionannot}) also provides the region's level
and variable annotation. 
Henceforth, we consider terms with fully annotated regions.
